Stochastic Blockmodels for Relational Event Data

### 2012-05-29 docs/submission.tex
Add Eckmann example figs to paper.

\documentclass{article}
\usepackage{nips12submit_e,times}
\usepackage{tikz}
\tikzstyle{vertex}=[auto=left,circle,fill=black!25,minimum size=20pt,inner sep=0pt]
\usepackage{verbatim}
\usepackage{url}
\usepackage{graphicx, color}
\usepackage{amsmath}
\usepackage{subfigure}

@@

\begin{figure}
\centering 
\subfigure[Dynamic network data]
{
\begin{tikzpicture}[scale=.8,transform shape,thick]
  \node[vertex] (n1) at (0,0)  {1};
  \node[vertex] (n2) at (2,2)  {2};
  \node[vertex] (n3) at (4,0) {3};
  \node[vertex] (n4) at (2,-2) {4};
\draw[-stealth] (.4,.4) -- (1.6,1.6);
\draw[-stealth] (2.4,1.6) -- (3.6,.4);
\node at (.7,1.1) {$t_1$};
\node at (3.3,1.1) {$t_2$};
\end{tikzpicture}
}
\subfigure[Intensities for two dyads]
{
\begin{tikzpicture}[scale=1,transform shape,thick]
\draw[-stealth] (0,0) -- (0,3);
\draw[-stealth] (0,0) -- (5,0);

% First timeline
\draw (0,2) -- (1,2);
\draw[fill=black] (1,2) circle (0.75mm);
\draw (1,1.7) circle (0.75mm);
\draw (1.08,1.7) -- (3,1.7);
\draw[fill=black] (3,1.7) circle (0.75mm);
\draw (3,1.7) circle (0.75mm);
\draw (3,2.2) circle (0.75mm);
\draw[-stealth] (3.05,2.2) -- (5,2.2);

% second timeline
\draw[dashed] (0,1) -- (3,1);
\draw[fill=black] (3,1) circle (0.75mm);
\draw (3,.75) circle (0.75mm);
\draw[dashed, -stealth] (3.08,.75) -- (5,.75);

% labels
\node at (-.5,2) {$\lambda_{1,2}$};
\node at (-.5,1) {$\lambda_{3,4}$};
\node at (1,-.3) {$t_1$};
\node at (3,-.3) {$t_2$};
\end{tikzpicture}
}
\caption[]{Illustration of event data and the assumptions of the model.  Left: An sequence of two events among four nodes: (1,2) occurs at time $t_1$ and (2,3) occurs at time $t_2$.  Right: The intensity functions $\lambda_{1,2}(t)$ (solid) and $\lambda_{34}(t)$ (dotted).  Note $\lambda_{34}$ can only change when events occur that involve either node 3 or node 4.\footnotemark }
\label{fig:example}
\end{figure}

@@

We aim to learn about the dynamics between subsets of nodes.   To facilitate this, we assume each node $i$ is assigned a latent cluster $z_i$ and use a log linear model for the intensity functions
\begin{align*}
\log \lambda_{ij}(t | \mathcal{A}_t,\mathbf{\beta},\mathbf{z}) &= \boldsymbol{\beta}'_{z_i,z_j} \mathbf{s}(t,i,j,\mathcal{A}_t)
%z_i &\sim \mbox{CRP}(\alpha) &
%\boldsymbol{\beta}_{k,l} &\sim %\mbox{N}_p(\boldsymbol{\mu},\boldsymbol{\sigma}^2I)
\end{align*}
where for each pair of clusters $(k,l)$ we have a vector of parameters $\boldsymbol{\beta}_{k,l} \sim \mbox{N}_p(\boldsymbol{\mu},\boldsymbol{\sigma}^2I)$ that corresponds to the vector of $P$ statistics $\mathbf{s}(t,i,j,\mathcal{A}_t)$ computed from the previous history $\mathcal{A}_t$.  Thus, our model for the rate of $(i,j)$ events has the same parameters as for other dyads occurring between group $z_i$ and $z_j$.

@@

\begin{figure}[t]
\includegraphics[width=3in]{../figs/eckmann-small/params-estimates}
\fbox{\rule[.5cm]{0cm}{5cm} \rule[.5cm]{5cm}{0cm}}
\caption{Left: Posterior credible intervals for parameter estimates from fitting the $K=2$ model to the Eckmann subset.  Right: Posterior distribution of $K$ when fitting the model and allowing for flexible $K$.}
\label{fig:posteriorparams}
\end{figure}

>>>>

\documentclass{article}
\usepackage{nips12submit_e,times}
\usepackage{tikz}
\tikzstyle{vertex}=[auto=left,circle,fill=black!25,minimum size=20pt,inner sep=0pt]
\usepackage{verbatim}
\usepackage{url}
\usepackage{graphicx, color}
\usepackage{amsmath}
\usepackage{caption}
\usepackage{subcaption}

@@

% \begin{figure}
% \centering 
% \begin{subfigure}
% \begin{tikzpicture}[scale=.8,transform shape,thick]
%   \node[vertex] (n1) at (0,0)  {1};
%   \node[vertex] (n2) at (2,2)  {2};
%   \node[vertex] (n3) at (4,0) {3};
%   \node[vertex] (n4) at (2,-2) {4};
% \draw[-stealth] (.4,.4) -- (1.6,1.6);
% \draw[-stealth] (2.4,1.6) -- (3.6,.4);
% \node at (.7,1.1) {$t_1$};
% \node at (3.3,1.1) {$t_2$};
% \end{tikzpicture}
% \caption{Dynamic network data}
% \end{subfigure}
% \begin{subfigure}
% \begin{tikzpicture}[scale=1,transform shape,thick]
% \draw[-stealth] (0,0) -- (0,3);
% \draw[-stealth] (0,0) -- (5,0);

% % First timeline
% \draw (0,2) -- (1,2);
% \draw[fill=black] (1,2) circle (0.75mm);
% \draw (1,1.7) circle (0.75mm);
% \draw (1.08,1.7) -- (3,1.7);
% \draw[fill=black] (3,1.7) circle (0.75mm);
% \draw (3,1.7) circle (0.75mm);
% \draw (3,2.2) circle (0.75mm);
% \draw[-stealth] (3.05,2.2) -- (5,2.2);

% % second timeline
% \draw[dashed] (0,1) -- (3,1);
% \draw[fill=black] (3,1) circle (0.75mm);
% \draw (3,.75) circle (0.75mm);
% \draw[dashed, -stealth] (3.08,.75) -- (5,.75);

% % labels
% \node at (-.5,2) {$\lambda_{1,2}$};
% \node at (-.5,1) {$\lambda_{3,4}$};
% \node at (1,-.3) {$t_1$};
% \node at (3,-.3) {$t_2$};
% \end{tikzpicture}
% \caption{Intensities for two dyads}
% \end{subfigure}
% \caption[]{Illustration of event data and the assumptions of the model.  Left: An sequence of two events among four nodes: (1,2) occurs at time $t_1$ and (2,3) occurs at time $t_2$.  Right: The intensity functions $\lambda_{1,2}(t)$ (solid) and $\lambda_{34}(t)$ (dotted).  Note $\lambda_{34}$ can only change when events occur that involve either node 3 or node 4.\footnotemark }
% \label{fig:example}
% \end{figure}

@@

We aim to learn about the dynamics between subsets of nodes.   To facilitate this, we assume each node $i$ is assigned a latent cluster $z_i$ and use a log linear model for the intensity functions
\begin{align*}
\log \lambda_{ij}(t | \mathcal{A}_t,\mathbf{\beta},\mathbf{z}) &= \boldsymbol{\beta}'_{z_i,z_j} \mathbf{s}(t,i,j,\mathcal{A}_t)
%z_i &\sim \mbox{CRP}(\alpha) &
%\boldsymbol{\beta}_{k,l} &\sim %\mbox{N}_p(\boldsymbol{\mu},\boldsymbol{\sigma}^2I)
\end{align*}
where for each pair of clusters $(k,l)$ we have a vector of parameters $\boldsymbol{\beta}_{k,l} \sim \mbox{N}_P(\boldsymbol{\mu},\boldsymbol{\sigma}^2I)$ that corresponds to the vector of $P$ statistics $\mathbf{s}(t,i,j,\mathcal{A}_t)$ computed from the previous history $\mathcal{A}_t$.  Thus, our model for the rate of $(i,j)$ events has the same parameters as for other dyads occurring between group $z_i$ and $z_j$.

@@

\begin{figure}[t]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics[scale=.5]{../figs/eckmann-small/example-obs-stats}
%\caption{Observed counts}
\end{subfigure}
\qquad
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics[scale=.5]{../figs/eckmann-small/example-estimates}
%\caption{Samples of $\beta_{p,k,l}$}
\end{subfigure}

%\fbox{\rule[.5cm]{0cm}{5cm} \rule[.5cm]{5cm}{0cm}}
\caption{%Observed counts versus model estimates from Eckmann email data.
 Left: Observed statistics across all dyadic email interactions from members of group 1 to group 3 (178 total events in total) and from members of group 3 to group 1 (179 events in total).  Right: Posterior samples of the corresponding $\beta_{p,1,3}$ and $\beta_{p,3,1}$.  In addition to finding differences in mean activity for these dyads, the model learns that events from 3 to 1 have a higher propensity for \texttt{ab-by} transitions.}
\label{fig:posteriorparams}
\end{figure}
